2.3 Prob Model for PCR Positivity Estimation

### 2024-01-24 2_scripts/2.3 Prob Model for PCR pos.qmd
Finish the model to predict 3-m PCR+

```{r}

```

>>>>

Compared to model 1, model 2 is predicting  "ever predicted" rather than "infected in past three months" as it does in model 1. The only one we can apply to the APL data is the model 1.

#### 2.1 Prepare for the dataset
```{r}
#those tested within last 90 days
summary(APL_PCR90$DAYS_SINCE_NAAT) # Model 1 is built upon this dataset

# Add an indicator for those who had PCR test in past 90 days
RFD4682_e$PCR90<-rep(0, length(RFD4682_e$order_ID))
RFD4682_e$PCR90[RFD4682_e$order_ID %in% APL_PCR90$order_ID]<-1
table(RFD4682_e$PCR90)

# Clean the gender variable (level unknown)
table(RFD4682_e$GENDER)
APL_Clean<-RFD4682_e
APL_Clean$GENDER[APL_Clean$GENDER=="Unknown"]<-NA

# Recode NAAT as 0/1/9
APL_Clean$NAAT[APL_Clean$NAAT_STATUS=="Positive"]<-1
APL_Clean$NAAT[APL_Clean$NAAT_STATUS=="Negative"]<-0
APL_Clean$NAAT[APL_Clean$NAAT_STATUS=="No Prior NAT"]<-9
APL_Clean$NAAT<-as.numeric(APL_Clean$NAAT)
table(APL_Clean$NAAT_STATUS, APL_Clean$NAAT, useNA = 'ifany')

# Apply model 1 to APL data (full)
APL_Clean$NAAT_p<-predict(plm01, newdata = APL_Clean, type = 'response')

# Create NAAT_m, using real NAAT results for those tested <=90 days
APL_Clean$NAAT_m <- APL_Clean$NAAT_p
APL_Clean$NAAT_m[APL_Clean$PCR90==1 & APL_Clean$NAAT_STATUS=="Positive"]<-1
APL_Clean$NAAT_m[APL_Clean$PCR90==1 & APL_Clean$NAAT_STATUS=="Negative"]<-0

summary(APL_Clean$NAAT_m)
table(APL_Clean$NAAT_m==1) # number matches
table(APL_Clean$NAAT_m==0)
```

#### 2.2 Based on new data, making plot seperately

```{r}
# Calculate positivity for each month

# tb1-the data (PCR<90) used to build model
tb1<-APL_Clean[APL_Clean$PCR90==1,] %>%
  group_by(Col_Month) %>%
  summarise(n = n() , pos= mean(NAAT_m))
tb1$PCR90<-rep(1, length(tb1$Col_Month))

# tb2-we applied the model to the rest of APL (PCR>90) 
tb2<-APL_Clean[APL_Clean$PCR90==0,] %>%
  group_by(Col_Month) %>%
  summarise(n = n() , pos= mean(NAAT_m, na.rm=TRUE))
tb2$PCR90<-rep(0, length(tb2$Col_Month))

# Merge the data to take a look
tb3<-merge(tb1, tb2, by = "Col_Month", all.x = T, all.y = T)
tb3$Col_Month<-as.yearmon(tb3$Col_Month)
tb3<-tb3[order(tb3$Col_Month),]

# Compare the difference
# Paired t-test
t.test(tb3$pos.x[18:27], tb3$pos.y[18:27], paired=1)
# p-value=0.2761, no sig diff between observed (<90D) and the rest
```
#### The plots
```{r}
APL_Pmelt<-rbind.data.frame(select(tb1, Col_Month, pos, PCR90), 
                            select(tb2[18:27,], Col_Month, pos, PCR90))
APL_Pmelt$PCR90[APL_Pmelt$PCR90==0]<-"Predicted"
APL_Pmelt$PCR90[APL_Pmelt$PCR90==1]<-"Observed"

APL_Pmelt %>%
  ggplot(aes(x=Col_Month, y=pos, group=PCR90, colour=PCR90)) +
  geom_point() +
  geom_line()  +
  geom_hline(yintercept=0, colour='blue', linetype=2) +
  theme_bw() 
```
